### 2018-05-17 Paper/method.tex
..

In this section, we outline the underlying framework for \emph{adaptively}
constructing a reduced-order surrogate (ROS) using sensitivity analysis.  The
proposed methodology is described as adaptive since the ROS is constructed only
in situations where it is expected to yield computational dividend as discussed
further below.  The term reduced-order implies that the surrogate is
constructed in a subspace that sufficiently captures the uncertainty in the
model output.~\alennote{The next sentence seems unneeded, and 
might be confusing; consider removing it, or clarify that in 
the case of PCE, we construct PC bases as multivariate polynomials
on a parameter space with reduced dimension.}
In the context of PC surrogates used in this work, it implies
that the dimensionality of the polynomial basis functions is effectively
reduced.

@@

In situations where $\nabla_{\bm{\theta}}G_k$ cannot be evaluated analytically,
we could use finite-difference to approximate its value as shown in
Eq.~\ref{eq:partial}. However, as discussed earlier in section 2.1,
~\alennote{Don't hard code section numbers, use labels, and refer to them.}
estimating $\bm{\mu}$  using $M$ samples in the full parameter space requires
model evaluations at $(N_p+1)M$ samples, compared to only $M$ samples in case
the derivative is available analytically. Hence, computational effort when
using finite differences is expected to increase by a factor of $(N_p+1)$.
Another possible approach, suitable in certain cases, is the use of
adjoint-based gradient computation. In such cases, each gradient computation,
which is more efficient than finite-differences---each gradient
evaluation requires a forward model solve and adjoint solve---but 
requires the availability of adjoint solvers.
\alennote{I will put some references there ...}

@@

The available set of model evaluations are utlized and further enhanced during
parameter screening as discussed earlier.  At the end of screening, the set of
active indices, $\mathcal{I}_\text{active}$, is used to evaluate $\alpha$ in the
following equation: 
\be
\alpha = \frac{|\mathcal{I}_\text{active}|}{N_p},
\label{eq:alpha}
\ee
\alennote{I noticed in the latex code the use of the \textbf{noindent} command;
please remove those, we will use latext setting to formation paragraphs.} where
$|\mathcal{I}_\text{active}|$ denotes the cardinality of
$\mathcal{I}_\text{active}$.  Scope for dimension-reduction increases as
$\alpha$ decreases.  Hence, if $\alpha$ is considered to be small and
computational gains are expected owing to dimension reduction, the ROS is
constructed and verified for accuracy using a combination of model evaluations
used for screening and those associated with the validation test suite.  On the
other hand, if $\alpha$ is considerably large (~\alennote{You mean close to
one?}), the set of inputs required for screening are updated as needed, and a
new set of $n_0$ samples and corresponding model evaluations are generated. The
FSS is re-constructed using the enriched set of evaluations and the
aforementioned analysis is repeated as illustrated in the flow-diagram
in Figure~\ref{fig:flow} that shows the overall 
parameter screening and surrogate model construction method.

\textbf{Validation metrics.}
To assess convergence of the resulting surrogate, one could estimate the
leave-one-out cross validation error as follows:
\be
\epsilon_{\mbox{\tiny LOO}} = 
\frac{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - G^{\mbox{\tiny {PC}\textbackslash i}}(\bm{\xi(\theta}_i))\right)^2}
{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - \widetilde{\mu}\right)^2}
\label{eq:loo}
\ee
where $N_l$ is the number of training points, 
$\widetilde{\mu}~=~\frac{1}{N_l}\sum\limits_{i=1}^{N_l} G(\bm{\theta}_i)$
is the sample mean of the model response, and $ G^{\mbox{\tiny
{PC}\textbackslash i}}$ is the PC surrogate constructed using all but the
$i^{\mbox{\tiny{th}}}$ model realization.  From~\eqref{eq:loo}, it appears
that $N_l$ PCEs are needed to evaluate $\epsilon_{\mbox{\tiny LOO}}$.  However,
in practice a modified formulation for $\epsilon_{\mbox{\tiny
LOO}}$~\cite{Blatman:2009}, independent of $G^{\mbox{\tiny {PC}\textbackslash
i}}$ is used.~\alennote{What is that modified formulation? Need to specify.}
Accuracy of the surrogate could be assessed by evaluating the
relative $\ell_2$-norm of the difference in predictions between the model and the
surrogate ($\epsilon_{\mbox{\tiny{L-2}}}$), as follows:
\be
\epsilon_{\mbox{\tiny{L-2}}} = \frac{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i) - 
G^{\mbox{\tiny {PC}}}(\bm{\xi(\theta}_i))\right)^2\right]^{\frac{1}{2}}}
{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i)\right)^2\right]^{\frac{1}{2}}}
\label{eq:l2}
\ee
where $N_v$ is the number of sampling points in the full parameter space at
which model evaluations are available. 
\alennote{You mean to do L2 error study in 
addition to the leave-one-out error studies?}
Accuracy of the surrogate could be
further investigated by comparing probability density functions (PDFs) of the
model output based on model evaluations in the full parameter space and
surrogate predictions corresponding to a large number of samples.  
\alennote{I don't think you can construct the PDF in full parameter 
space, in real applications, as you need a large number of 
expensive model evaluations. Maybe we should say that  
we compare the PDF of the surrogate to a histogram of true model
evaluations?}

>>>>

In this section, we outline the underlying framework for \emph{adaptively}
constructing a reduced-order surrogate (ROS) using sensitivity analysis.  The
proposed methodology is described as adaptive since the ROS is constructed only
in situations where it is expected to yield computational dividend as discussed
further below.  The term reduced-order implies that the surrogate is
constructed in a subspace that sufficiently captures the uncertainty in the
model output. We begin by outlining an algorithm for parameter screening
to assess the importance of individual parameters for potential dimension
reduction and construction of an ROS. The overall adaptive framework 
which incorporates parameter screening as an integral step is thereafter
presented. Finally, we present validation metrics used for assessing the
convergence and accuracy of the ROS followed by a brief discussion on
salient features of the proposed framework.

@@

In situations where $\nabla_{\bm{\theta}}G_k$ cannot be evaluated analytically,
we could use finite-difference to approximate its value as shown in
Eq.~\ref{eq:partial}. However, as discussed earlier in~\ref{sub:dgsm},
estimating $\bm{\mu}$  using $M$ samples in the full parameter space requires
model evaluations at $(N_p+1)M$ samples, compared to only $M$ samples in case
the derivative is available analytically. Hence, computational effort when
using finite differences is expected to increase by a factor of $(N_p+1)$.
Another possible approach, suitable in certain cases, is the use of
adjoint-based gradient computation. In such cases, each gradient computation,
which is more efficient than finite-differences---each gradient
evaluation requires a forward model solve and adjoint solve---but 
requires the availability of adjoint solvers.
\alennote{I will put some references there ...}

@@

The available set of model evaluations are utlized and further enhanced during
parameter screening as discussed earlier.  At the end of screening, the set of
active indices, $\mathcal{I}_\text{active}$, is used to evaluate $\alpha$ in the
following equation: 
\be
\alpha = \frac{|\mathcal{I}_\text{active}|}{N_p},
\label{eq:alpha}
\ee
\alennote{I noticed in the latex code the use of the \textbf{noindent} command;
please remove those, we will use latext setting to formation paragraphs.} where
$|\mathcal{I}_\text{active}|$ denotes the cardinality of
$\mathcal{I}_\text{active}$.  Scope for dimension-reduction increases as
$\alpha$ decreases.  Hence, if $\alpha$ is considered to be small and
computational gains are expected owing to dimension reduction, the ROS is
constructed and verified for accuracy using a combination of model evaluations
used for screening and those associated with the validation test suite.  On the
other hand, if $\alpha$ is close to 1, the set of inputs required for screening are
 updated as needed, and a
new set of $n_0$ samples and corresponding model evaluations are generated. The
FSS is re-constructed using the enriched set of evaluations and the
aforementioned analysis is repeated as illustrated in the flow-diagram
in Figure~\ref{fig:flow} that shows the overall 
parameter screening and surrogate model construction method.

\textbf{Validation metrics.}
To assess convergence of the resulting surrogate, one could estimate the
leave-one-out cross validation error as follows:
\be
\epsilon_{\mbox{\tiny LOO}} = 
\frac{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - G^{\mbox{\tiny {PC}\textbackslash i}}(\bm{\xi(\theta}_i))\right)^2}
{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - \widetilde{\mu}\right)^2}
\label{eq:loo}
\ee
where $N_l$ is the number of training points, 
$\widetilde{\mu}~=~\frac{1}{N_l}\sum\limits_{i=1}^{N_l} G(\bm{\theta}_i)$
is the sample mean of the model response, and $ G^{\mbox{\tiny
{PC}\textbackslash i}}$ is the PC surrogate constructed using all but the
$i^{\mbox{\tiny{th}}}$ model realization.  From~\eqref{eq:loo}, it appears
that $N_l$ PCEs are needed to evaluate $\epsilon_{\mbox{\tiny LOO}}$.  However,
in practice a modified formulation for $\epsilon_{\mbox{\tiny
LOO}}$~\cite{Blatman:2009}, independent of $G^{\mbox{\tiny {PC}\textbackslash
i}}$ is used.~\alennote{What is that modified formulation? Need to specify.}
Accuracy of the surrogate could be assessed by evaluating the
relative $\ell_2$-norm of the difference in predictions between the model and the
surrogate ($\epsilon_{\mbox{\tiny{L-2}}}$), as follows:
\be
\epsilon_{\mbox{\tiny{L-2}}} = \frac{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i) - 
G^{\mbox{\tiny {PC}}}(\bm{\xi(\theta}_i))\right)^2\right]^{\frac{1}{2}}}
{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i)\right)^2\right]^{\frac{1}{2}}}
\label{eq:l2}
\ee
where $N_v$ is the number of sampling points in the full parameter space at
which model evaluations are available. 
\alennote{You mean to do L2 error study in 
addition to the leave-one-out error studies?}
Accuracy of the surrogate could be
further investigated by comparing probability density functions (PDFs) of the
model output based on model evaluations in the full parameter space and
the ROS predictions corresponding to a large number of samples (say, 10$^6$). 
However, in realistic problems involving complex, compute-intensive 
simulations, the PDF based on model evaluations would be intractable.
A practical alternative would be to compare a histogram
based on sparse model evaluations with the surrogate-based PDF
in order to gain some insight into the quality of the comparison.
